Equalizer Programming Guide

### 2008-10-27 Developer/ProgrammingGuide/paper.tex
Programming Guide: Add DPlex documentation
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.3.14 for Equalizer 0.5.5+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add \sref{sStatistics}, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10-19, 2008 & add \aref{aFileFormat}\\
    1.3.4   & July 3, 2008 & Update \sref{sStatistics}\\
    1.3.5   & July 7, 2008 & Update after namespace naming cleanup\\
    1.3.6   & July 20, 2008 & Add accum and samples to \aref{aFileFormat}, add \sref{sDrawableConfig}\\
    1.3.7   & Aug 22, 2008 & Update \sref{sEqplyWIndow} with OSWindow changes\\
    1.3.8   & Aug 25, 2008 & Update \sref{sEventHandling} with OSWindow changes\\
    1.3.9   & Aug 29, 2008 & Update \sref{sEventHandling}: event handling cleanup\\
    1.3.10  & Oct 7, 2008  & Updated to reflect 2D pixel compounds\\
    1.3.11  & Oct 13, 2008 & Updated to reflect eq::net::Object cleanup\\
    1.3.12  & Oct 16, 2008 & Added \sref{sView}\\
    1.3.13  & Oct 17, 2008 & Minor update in \sref{sMessagePump}\\
    1.3.14  & Oct 21, 2008 & Improved \sref{sConnectionDescription}\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/2D.pdf}
  {\caption{\small A 2D compound}}
\end{wrapfigure}
2D or sort-first decomposes the rendering in screen-space, that is, each
contributing rendering unit processes a tile of the final view. The
recomposition simply assembles the tiles side-by-side on the destination
view.

The advantage of this mode is a low, constant IO overhead for the pixel
transfers, since only color information has to be transmitted. The upper
limit is the amount of pixel data for the destination view.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/DB.pdf}
  {\caption{\small A database compound}}
\end{wrapfigure}
DB or sort-last\footnote{3D model courtesy of AVS, USA.} decomposes the
rendered database so that all rendering units process a part of the
scene in parallel.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/EYE.pdf}
  {\caption{\small A stereo compound}}
\end{wrapfigure}
Stereo compounds\footnote{3D model courtesy of Stereolithography Archive
  at Clemson University.} assign each eye pass to individual rendering
units. The resulting images are copied to the appropriate stereo
buffer. This modes supports a variety of stereo modes, including active
(quad-buffered) stereo, anaglyphic stereo and auto-stereo displays with
multiple eye passes.

Due to the frame consistency between the eye views, this modes scales
very well. The IO requirements for pixel transfer are small and
constant.

@@

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frustra.pdf}
  {\caption{\small\label{fFrustra}Wall and projection parameters}}
\end{wrapfigure}
The frustum description is inherited by the children, therefore the
frustum is typically defined on the top-most compound.

@@

A wall is completely defined by the bottom-left, bottom-right and
top-left coordinates relative to the origin.

@@

The tasks are executed by the node and pipe threads in parallel. The
parent's initialization methods are always executed before any child
initialization method. This parallelization is necessary to allow a
speedy startup of the configuration on large-scale graphics clusters. On
the other hand, it means that initialization functions are called even
if the parent's initialization has failed.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/configInit.pdf}
  {\caption{\small\label{fConfigInit}Config Initialization Sequence}}
\end{wrapfigure}
If the configuration was initialized correctly, the configuration tries
to set up a tracking device for head tracking. Equalizer does not
provide extensive support for tracking devices, as this is an orthogonal
problem to parallel rendering. Tracking device support has already been
solved by a number of implementations\footnote{VRCO Trackd VRPN, etc.},
which can easily be integrated with Equalizer. The example code in
\textsf{eqPly} provides a reference implementation for the integration
of such a tracking library. \sref{sTracking} provides more background on
head tracking.

@@

All window-system specific functionality is implemented by a
specialization of \textsf{eq::OS\-Window}. The \textsf{OSWindow} class
defines the minimal interface to be implemented for a new window
system. Each \textsf{Window} uses one \textsf{OSWindow} during
execution. This separation allows an easy implementation and adaption to
another window system or application.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/osWindow.pdf}
  {\caption{\small\label{fOSWindow}eq::OSWindow UML class hierarchy}}
\end{wrapfigure}
The \textsf{OSWindow} initialization takes into account various
attributes set in the configuration file. Attributes include the size of
the various frame buffer planes (color, alpha, depth, stencil) as well
as other framebuffer attributes, such as quad-buffered stereo,
doublebuffering, fullscreen mode and window decorations. Some of the
attributes, such as stereo, doublebuffer and stencil can be set to
\textsf{eq::AUTO}, in which case the Equalizer default implementation
will test for their availability and enable them if possible.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}\includegraphics[width=.618\textwidth]{images/render.pdf}
  {\caption{\small\label{fRender}Main Render Loop}}
  \vspace{-5ex}
\end{wrapfigure}
The model data is spatially organized in an 3-dimensional
kD-tree\footnote{See also \link{http://en.wikipedia.org/wiki/Kd-tree}}
for efficient view frustum culling. When the model is loaded by
\textsf{Node::config\-Init}, it is preprocessed into the kD-tree During
this preprocessing step, each node of the tree gets a database range
assigned. The root node has the range [0, 1], its left child [0, 0.5]
and its right child [0.5, 1], and so on for all nodes in the tree. The
preprocessed model is saved in a binary format for accelerating
subsequent loading.

The rendering loop maintains a list of candidates to render, which
initially contains the root node. Each candidate of this list is tested
for full visiblity against the frustum and range, and rendered if
visible. It is dropped if it is fully invisible or fully out of
range. If it is partially visible or partially in range, the children of
the node are added to the candidate list. \fig{fRender} shows a flow
chart of the rendering algorithm, which performs efficient view frustum
and range culling.

@@

The purpose of the \textsf{OSWindow} processing method,
\textsf{processEvent}, is to perform window-system specific event
processing. For example, \textsf{AGLWindow::processEvent} calls
\textsf{aglUpdateContext} whenever a resize event is received. For
further processing, the \textsf{Event} is passed on to
\textsf{Window::processEvent}. This \textsf{Event} no longer contains
the native event.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/eventFilter.pdf}
  {\caption{\small\label{fEventProcessing}Event Processing}}
\end{wrapfigure}
\textsf{Window::processEvent} is responsible for either handling
the event locally, or for translating it into a generic
\textsf{ConfigEvent}. The config events are send to the application
thread using \textsf{Config::sendEvent}.

@@

The concrete implementation of these window interfaces is responsible
for setting up and de-initializing event handling.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/eventUML.pdf}
  {\caption{\small\label{fEventUML}UML Diagram for Event Handling Classes}}
\end{wrapfigure}
The glX event handler iterates over all windows of the registered pipe
to find the originating \textsf{GLXWindowIF}, which is then used to
process the event in a fashion similar to AGL. The
\textsf{GLXWindowEvent} passes the \textsf{XEvent} to
\textsf{GLXWindowIF::\-process\-Event}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/threads.pdf}
  {\caption{\small\label{fThreads}Threads within one node process}}
\end{wrapfigure}
The receiver and command thread are terminated when the application
stops network communications using \textsf{Client::exitLocal}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frameSync.pdf}
  {\caption{\small\label{fFrameSync}Per-Node Frame Synchronization}}
\end{wrapfigure}
The application has extended control over the task synchronization
during a frame. Upon \textsf{Config::startFrame}, Equalizer invokes the
\textsf{frameStart} task methods of the various entities. The entities
unlock all their children by calling \textsf{startFrame}, e.g.,
\textsf{Node::frameStart} has to call \textsf{Node::startFrame} in order
to unlock the pipe threads. Note that certain \textsf{startFrame} calls,
e.g., \textsf{Window::startFrame}, are currently empty since the
synchronization is implicit due to the sequential execution within the
thread.

Each entity uses \textsf{waitFrame\-Started} to block on the parent's
\textsf{startFrame}, e.g., \textsf{Pipe::\-frame\-Start} calls
\textsf{Node::wait\-Frame\-Started} to wait for the corresponding
\textsf{Node::start\-Fra\-me}. This explicit synchronization allows to
update non-critical data before synchronizing with
\textsf{waitFrameStarted}, or after unlocking using
\textsf{start\-Fra\-me}. \fig{fFrameSync} illustrates this
synchronization model.

@@

The main idea behind direct send is to parallelize the costly
recomposition for database (sort-last) decomposition. With each
additional source channel, the amount of pixel data to be composited
grows linearly. When using the simple approach of compositing all frames
on the destination channel, this channel quickly becomes the bottleneck
in the system. Direct send distributes this workload evenly across all
source channels, and thereby keeps the compositing work per channel
constant.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-1ex}\includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\small\label{fDirectSend}Parallel Direct Send Compositing}}
  \vspace{-2ex}
\end{wrapfigure}
After performing the sort-last compositing, the color information is
transferred to the destination channel, similarly to a 2D (sort-first)
compound. The amount of pixel data for this part of the compositing
pipeline also approaches a constant value, i.e., the full frame buffer.

@@

The buffer defines the default frame buffer attachments read back by
output frames. The viewport restricts the rendering to the area relative
to the parent compound. The range restricts the database range, relative
to the parent. The pixel setting selects the pixel decomposition kernel,
relative to the parent. The eye pass select the mono, left or right eye
passes rendered by the compound. All these attributes are inherited by
the children of a compound. Viewport, range and pixel parameters are
cumulative.

@@

{\footnotesize\begin{lstlisting}
compound
{
    name     string
    channel  string   // where the compound's tasks are executed
    task     [ CLEAR DRAW READBACK ASSEMBLE ] // CULL later
    buffer   [ COLOR DEPTH ]         // default COLOR
    viewport [ viewport ]            // wrt parent compound, sort-first
    range    [ float float ]         // DB-range for sort-last
    pixel    [ int int ]             // pixel decomposition (step size)
    eye      [ CYCLOP LEFT RIGHT ]   // monoscopic or stereo view
    attributes
    {
        stereo_mode                  QUAD | ANAGLYPH    // default QUAD
        stereo_anaglyph_left_mask    [ RED GREEN BLUE ] // default red
        stereo_anaglyph_right_mask   [ RED GREEN BLUE ] // df green blue
        update_FOV                   HORIZONTAL [ON] | VERTICAL | OFF
    }

>>>>

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.3.15 for Equalizer 0.5.5+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add \sref{sStatistics}, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10-19, 2008 & add \aref{aFileFormat}\\
    1.3.4   & July 3, 2008 & Update \sref{sStatistics}\\
    1.3.5   & July 7, 2008 & Update after namespace naming cleanup\\
    1.3.6   & July 20, 2008 & Add accum and samples to \aref{aFileFormat}, add \sref{sDrawableConfig}\\
    1.3.7   & Aug 22, 2008 & Update \sref{sEqplyWIndow} with OSWindow changes\\
    1.3.8   & Aug 25, 2008 & Update \sref{sEventHandling} with OSWindow changes\\
    1.3.9   & Aug 29, 2008 & Update \sref{sEventHandling}: event handling cleanup\\
    1.3.10  & Oct 7, 2008  & Updated to reflect 2D pixel compounds\\
    1.3.11  & Oct 13, 2008 & Updated to reflect eq::net::Object cleanup\\
    1.3.12  & Oct 16, 2008 & Added \sref{sView}\\
    1.3.13  & Oct 17, 2008 & Minor update in \sref{sMessagePump}\\
    1.3.14  & Oct 21, 2008 & Improved \sref{sConnectionDescription}\\
    1.3.15  & Oct 27, 2008 & Added \sref{sDPlex}\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

2D decomposes the rendering in screen-space, that is, each contributing
rendering unit processes a tile of the final view. The recomposition
simply assembles the tiles side-by-side on the destination view. This
mode is also known as sort-first or SFR.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/2D.pdf}
  {\caption{\small A 2D compound}}
\end{wrapfigure}
The advantage of this mode is a low, constant IO overhead for the pixel
transfers, since only color information has to be transmitted. The upper
limit is the amount of pixel data for the destination view.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/DB.pdf}
  {\caption{\small A database compound}}
\end{wrapfigure}
DB\footnote{3D model courtesy of AVS, USA.} decomposes the
rendered database so that all rendering units process a part of the
scene in parallel. This mode is also known as sort-last.

@@

Stereo compounds\footnote{3D model courtesy of Stereolithography Archive
  at Clemson University.} assign each eye pass to individual rendering
units. The resulting images are copied to the appropriate stereo
buffer. This mode supports a variety of stereo modes, including active
(quad-buffered) stereo, anaglyphic stereo and auto-stereo displays with
multiple eye passes.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/EYE.pdf}
  {\caption{\small A Stereo compound}}
\end{wrapfigure}
Due to the frame consistency between the eye views, this modes scales
very well. The IO requirements for pixel transfer are small and
constant.

@@

\subsection{\label{sDPlex}DPlex Compounds}

DPlex compounds\footnote{3D model courtesy of Cyberware, Inc.} assign
full, alternating frames to individual rendering units. The resulting
images are copied to the destination channel, and Equalizer
load-balancing is used to ensure a steady framerate on the destination
window. This mode is also known as time-multiplex or AFR.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/DPlex.pdf}
  {\caption{\small A DPlex compound}}
\end{wrapfigure}

Due to the frame consistency between consecutive frames, this mode scales
very well. The IO requirements for pixel transfer are small and
constant.

DPlex requires a latency of at least n frames. This increased latency
might be disturbing to the user, but it is often compensated by the
higher frame rate. The frame rate typically increases linearly with the
number of source channels, and therefore linearly with the latency.

DPlex compounds in Equalizer are configured using the \textsf{period}
and \textsf{phase} parameter, limiting each child to render a subset of
the frames. Each child compound uses an output frame, which is connected
to an input frame on the destination channel. The destination channel
uses a DPlex load-balancer to smoothen the framerate.

@@

The frustum description is inherited by the children, therefore the
frustum is typically defined on the top-most compound.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frustra.pdf}
  {\caption{\small\label{fFrustra}Wall and projection parameters}}
\end{wrapfigure}
A wall is completely defined by the bottom-left, bottom-right and
top-left coordinates relative to the origin.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/configInit.pdf}
  {\caption{\small\label{fConfigInit}Config Initialization Sequence}}
\end{wrapfigure}
The tasks are executed by the node and pipe threads in parallel. The
parent's initialization methods are always executed before any child
initialization method. This parallelization is necessary to allow a
speedy startup of the configuration on large-scale graphics clusters. On
the other hand, it means that initialization functions are called even
if the parent's initialization has failed.

@@

If the configuration was initialized correctly, the configuration tries
to set up a tracking device for head tracking. Equalizer does not
provide extensive support for tracking devices, as this is an orthogonal
problem to parallel rendering. Tracking device support has already been
solved by a number of implementations\footnote{VRCO Trackd VRPN, etc.},
which can easily be integrated with Equalizer. The example code in
\textsf{eqPly} provides a reference implementation for the integration
of such a tracking library. \sref{sTracking} provides more background on
head tracking.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/osWindow.pdf}
  {\caption{\small\label{fOSWindow}eq::OSWindow UML class hierarchy}}
\end{wrapfigure}
All window-system specific functionality is implemented by a
specialization of \textsf{eq::OS\-Window}. The \textsf{OSWindow} class
defines the minimal interface to be implemented for a new window
system. Each \textsf{Window} uses one \textsf{OSWindow} during
execution. This separation allows an easy implementation and adaption to
another window system or application.

@@

The \textsf{OSWindow} initialization takes into account various
attributes set in the configuration file. Attributes include the size of
the various frame buffer planes (color, alpha, depth, stencil) as well
as other framebuffer attributes, such as quad-buffered stereo,
doublebuffering, fullscreen mode and window decorations. Some of the
attributes, such as stereo, doublebuffer and stencil can be set to
\textsf{eq::AUTO}, in which case the Equalizer default implementation
will test for their availability and enable them if possible.

@@

The model data is spatially organized in an 3-dimensional
kD-tree\footnote{See also \link{http://en.wikipedia.org/wiki/Kd-tree}}
for efficient view frustum culling. When the model is loaded by
\textsf{Node::config\-Init}, it is preprocessed into the kD-tree During
this preprocessing step, each node of the tree gets a database range
assigned. The root node has the range [0, 1], its left child [0, 0.5]
and its right child [0.5, 1], and so on for all nodes in the tree. The
preprocessed model is saved in a binary format for accelerating
subsequent loading.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/render.pdf}
  {\caption{\small\label{fRender}Main Render Loop}}
\end{wrapfigure}
The rendering loop maintains a list of candidates to render, which
initially contains the root node. Each candidate of this list is tested
for full visiblity against the frustum and range, and rendered if
visible. It is dropped if it is fully invisible or fully out of
range. If it is partially visible or partially in range, the children of
the node are added to the candidate list. \fig{fRender} shows a flow
chart of the rendering algorithm, which performs efficient view frustum
and range culling.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/eventFilter.pdf}
  {\caption{\small\label{fEventProcessing}Event Processing}}
\end{wrapfigure}
The purpose of the \textsf{OSWindow} processing method,
\textsf{processEvent}, is to perform window-system specific event
processing. For example, \textsf{AGLWindow::processEvent} calls
\textsf{aglUpdateContext} whenever a resize event is received. For
further processing, the \textsf{Event} is passed on to
\textsf{Window::processEvent}. This \textsf{Event} no longer contains
the native event.

\textsf{Window::processEvent} is responsible for either handling
the event locally, or for translating it into a generic
\textsf{ConfigEvent}. The config events are send to the application
thread using \textsf{Config::sendEvent}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/eventUML.pdf}
  {\caption{\small\label{fEventUML}UML Diagram for Event Handling Classes}}
\end{wrapfigure}
The concrete implementation of these window interfaces is responsible
for setting up and de-initializing event handling.

@@

The glX event handler iterates over all windows of the registered pipe
to find the originating \textsf{GLXWindowIF}, which is then used to
process the event in a fashion similar to AGL. The
\textsf{GLXWindowEvent} passes the \textsf{XEvent} to
\textsf{GLXWindowIF::\-process\-Event}.

@@

The receiver and command thread are terminated when the application
stops network communications using \textsf{Client::exitLocal}.

@@

The application has extended control over the task synchronization
during a frame. Upon \textsf{Config::startFrame}, Equalizer invokes the
\textsf{frameStart} task methods of the various entities. The entities
unlock all their children by calling \textsf{startFrame}, e.g.,
\textsf{Node::frameStart} has to call \textsf{Node::startFrame} in order
to unlock the pipe threads. Note that certain \textsf{startFrame} calls,
e.g., \textsf{Window::startFrame}, are currently empty since the
synchronization is implicit due to the sequential execution within the
thread.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frameSync.pdf}
  {\caption{\small\label{fFrameSync}Per-Node Frame Synchronization}}
\end{wrapfigure}
Each entity uses \textsf{waitFrame\-Started} to block on the parent's
\textsf{startFrame}, e.g., \textsf{Pipe::\-frame\-Start} calls
\textsf{Node::wait\-Frame\-Started} to wait for the corresponding
\textsf{Node::start\-Fra\-me}. This explicit synchronization allows to
update non-critical data before synchronizing with
\textsf{waitFrameStarted}, or after unlocking using
\textsf{start\-Fra\-me}. \fig{fFrameSync} illustrates this
synchronization model.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-1ex}\includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\small\label{fDirectSend}Parallel Direct Send Compositing}}
\end{wrapfigure}
The main idea behind direct send is to parallelize the costly
recomposition for database (sort-last) decomposition. With each
additional source channel, the amount of pixel data to be composited
grows linearly. When using the simple approach of compositing all frames
on the destination channel, this channel quickly becomes the bottleneck
in the system. Direct send distributes this workload evenly across all
source channels, and thereby keeps the compositing work per channel
constant.

@@

After performing the sort-last compositing, the color information is
transferred to the destination channel, similarly to a 2D (sort-first)
compound. The amount of pixel data for this part of the compositing
pipeline also approaches a constant value, i.e., the full frame buffer.

@@

The buffer defines the default frame buffer attachments read back by
output frames. The viewport restricts the rendering to the area relative
to the parent compound. The range restricts the database range, relative
to the parent. The pixel setting selects the pixel decomposition kernel,
relative to the parent. The eye pass select the mono, left or right eye
passes rendered by the compound. The DPlex period defines that
$\frac{1}{period}$ frames are rendered, and the phase defines when in the period
the rendering starts. All these attributes are inherited by the children
of a compound. Viewport, range, pixel and period parameters are
cumulative.

A load-balancer is used to automatically optimize the decomposition. A
2D, horizontal or vertical load-balancer adapts the viewport of all
direct children of the compound each frame. A DB load-balancer adapts
the range of all direct children. A DPlex load-balancer smoothens the
framerate of the compound's window to produce a steady output framerate.

@@

{\footnotesize\begin{lstlisting}
compound
{
    name     string
    channel  string   // where the compound's tasks are executed
    task     [ CLEAR DRAW READBACK ASSEMBLE ] // CULL later
    buffer   [ COLOR DEPTH ]         // default COLOR
    viewport [ viewport ]            // wrt parent compound, sort-first
    range    [ float float ]         // DB-range for sort-last
    pixel    [ int int ]             // pixel decomposition (step size)
    eye      [ CYCLOP LEFT RIGHT ]   // monoscopic or stereo view
    period   int                     // DPlex period, every nth frame rendered
    phase    int                     // DPlex phase, when to start rendering

    loadBalancer
    {
        mode   2D | HORIZONTAL | VERTICAL | DB | DPLEX
    }

    attributes
    {
        stereo_mode                  QUAD | ANAGLYPH    // default QUAD
        stereo_anaglyph_left_mask    [ RED GREEN BLUE ] // default red
        stereo_anaglyph_right_mask   [ RED GREEN BLUE ] // df green blue
        update_FOV                   HORIZONTAL [ON] | VERTICAL | OFF
    }
